Complexity analysis and a fully polynomial-time approximation scheme for the college application problem

### 2022-06-28 paper/paper.tex
line edits; add affil info

\documentclass[11pt]{article} %try amsproc, amsart
%\usepackage{geometry}       \geometry{a4paper, margin=2.5cm}  
\usepackage{hyperref}       \hyperbaseurl{}    
                            \urlstyle{same}
\usepackage[ruled,linesnumbered]{algorithm2e}   \SetKwInOut{KwParams}{Parameters}
                                                \newcommand{\lIfElse}[3]{\lIf{#1}{#2 \textbf{else}~#3}}
\usepackage{stackengine}    \newcommand\xrowht[2][0]{\addstackgap[.5\dimexpr#2\relax]{\vphantom{#1}}}

@@

\title{Complexity analysis and a fully polynomial-time approximation scheme for the college application problem}
\author[1]{Max Kapur}
\author[2]{Sung-Pil Hong}
\affil[1,2]{Department of Industrial Engineering, Seoul National University}
\affil[1]{Email: maxkapur@gmail.com}
\date{\today}

@@

\begin{abstract}
The college application problem is a submodular optimization problem involving the maximization of the expected maximum value of a portfolio of random variables subject to a budget constraint. We introduce this problem, show that it is NP-complete, and provide two optimization algorithms based on dynamic programming. A modification of the second dynamic program yields a fully polynomial-time approximation scheme.
\end{abstract}

@@

\begin{proof}
(CAP $\in$ NP.) It suffices to show that the cost and value of a given $\mathcal{X} \subseteq \mathcal{C}$ can be computed in polynomial time. Given an oracle that multiplies integers in unit time, $v(\mathcal{X})$ can be computed in polynomial time using the following algorithm (which assumes that the elements of $\mathcal{X}$ are iterated in index order):
\filbreak
\RestyleAlgo{plain}
\LinesNotNumbered 
\begin{algorithm}[h]
\SetAlgoCaptionSeparator{}% no separator, default colon
\SetAlCapNameSty{}% no caption text
$v \gets 0$\;
\For{$j \in \mathcal{X}$}{
	$v \gets (1 - f_j)\,v + f_j t_j$\;
}
\Return{$v$}\;
\end{algorithm}

@@

\subsection{Dynamic program based on portfolio valuations}

@@

The details of the FPTAS are similar to those of Algorithm \ref{ellisDP2}. For integers $0 \leq j \leq m$ and $v \in \mathbb{Z}/2^P \cap \mathcal{V}'$, $\mathcal{W}[j, v]$ denotes the least expensive portfolio that uses only schools $\{ 1, \dots, j\}$ and has valuation of \emph{at least} $v$, if such a portfolio exists. Its cost is given by $G[j, v] = \sum_{j\in \mathcal{W}[j, v]} g_j$. The boundary conditions and recursion relation are identical to those above, except that $\Delta_j(v)$ is replaced with
\begin{align}
\Delta_j[v]
\begin{cases}
r\left[\frac{f_j}{1 - f_j} (t_j - v)\right], \quad & f_j < 1\\
\infty, &f_j = 1.
\end{cases} \label{clampeddeltajvdef}
\end{align}
By rounding the solution of \eqref{solvemeforvwjvminusj} down, we ensure that the true valuation of $\mathcal{W}[j, v]$ is \emph{at least} $v - \Delta$, and therefore that $G[j, v]$ meets its definition.

@@

Algorithms \ref{ellisDP1}, \ref{ellisDP2}, and \ref{ellisDP3} can be written using recursive functions instead of lookup tables. However, since each function references itself \emph{twice,} the function values at each index must be recorded in the computer's memory to take advantage of overlapping indices and guarantee the runtimes promised above. A discussion of these and other implementation issues is provided in Kapur (2022), along with a computational study demonstrating the efficiency of Algorithms \ref{ellisDP1} and \ref{ellisDP3}.

@@

This article has introduced a novel nonlinear knapsack problem called the college application problem. Our results suggest that the college application is a relatively easy instance of submodular maximization: Whereas $1 - 1/e$ is the highest approximation coefficient achievable by a polynomial-time algorithm for nondecreasing submodular maximization over a knapsack constraint (Nemhauser and Wolsey 1978), the FPTAS presented here induces a continuum of tight, polynomial-time approximation algorithms. This result suggests that though the objective function of the college application is nonlinear, it has similar regularity properties to a linear function. Further evidence for this statement is given by Kapur (2022), who proved the optimality of a greedy algorithm for maximizing the college application problem's objective function over a cardinality constraint.

We conjecture that other results associated with maximizing linear functions may hold with respect to the objective discussed here, such the existence of a PTAS for the multiple knapsack problem, and the optimality of the greedy algorithm a over polymatroidal constraint (Chekuri and Khanna 2005; Edmonds 1971).

>>>>

\documentclass[11pt]{article} %try amsproc, amsart
\usepackage{geometry}       \geometry{a4paper, margin=2.5cm}
\usepackage{setspace}       \setstretch{1.1}
\usepackage{hyperref}       \hyperbaseurl{}    
                            \urlstyle{same}
\usepackage[ruled,linesnumbered]{algorithm2e}   \SetKwInOut{KwParams}{Parameters}
                                                \newcommand{\lIfElse}[3]{\lIf{#1}{#2 \textbf{else}~#3}}
\usepackage{stackengine}    \newcommand\xrowht[2][0]{\addstackgap[.5\dimexpr#2\relax]{\vphantom{#1}}}

@@

\title{Complexity analysis and a fully polynomial-time approximation scheme for the college application problem}
\author[1]{Max Kapur}
\author[2]{Sung-Pil Hong}
\affil[1,2]{Department of Industrial Engineering, Seoul National University}
\date{\today}

@@

\begin{abstract}
The college application problem is a submodular optimization problem involving the maximization of the expected maximum value of a portfolio of random variables subject to a budget constraint. We introduce this problem, show that it is NP-complete, and provide two optimization algorithms based on dynamic programming. A modification of the second dynamic program yields a fully polynomial-time approximation scheme.

\textbf{Keywords:} submodular maximization, knapsack problems, approximation algorithms

\vfill
Correspondence may be addressed to Max Kapur.
\begin{itemize}
%\item[] Correspondence may be addressed to Max Kapur.

\item[] Email: \url{maxkapur@gmail.com}

\item[] Address: 39-411, 1 Gwanak-ro, Gwanak-gu, Seoul 08826, Republic of Korea
\end{itemize}

\end{abstract}

@@

\begin{proof}
(CAP $\in$ NP.) It suffices to show that the cost and value of a given $\mathcal{X} \subseteq \mathcal{C}$ can be computed in polynomial time. Given an oracle that multiplies integers in unit time, $v(\mathcal{X})$ can be computed in polynomial time using the following algorithm (which assumes that the elements of $\mathcal{X}$ are iterated in index order):

\RestyleAlgo{plain}
\LinesNotNumbered 
\begin{algorithm}[h]
\SetAlgoCaptionSeparator{}% no separator, default colon
\SetAlCapNameSty{}% no caption text
$v \gets 0$\;
\For{$j \in \mathcal{X}$}{
	$v \gets (1 - f_j)\,v + f_j t_j$\;
}
\Return{$v$}\;
\end{algorithm}

@@

\subsection{Dynamic program based on portfolio valuations} \label{dpbasedonportfoliovaluations}

@@

The details of the FPTAS are similar to those of Algorithm \ref{ellisDP2}. For integers $0 \leq j \leq m$ and $v \in \mathbb{Z}/2^P \cap \mathcal{V}'$, $\mathcal{W}[j, v]$ denotes the least expensive portfolio that uses only schools $\{ 1, \dots, j\}$ and has valuation of \emph{at least} $v$, if such a portfolio exists. Its cost is given by $G[j, v] = \sum_{j\in \mathcal{W}[j, v]} g_j$. The boundary conditions and recursion relation are identical to those given above (subsection \ref{dpbasedonportfoliovaluations}), except that $\Delta_j(v)$ is replaced with
\begin{align}
\Delta_j[v]
\begin{cases}
r\left[\frac{f_j}{1 - f_j} (t_j - v)\right], \quad & f_j < 1\\
\infty, &f_j = 1.
\end{cases} \label{clampeddeltajvdef}
\end{align}
Here we have rounded the solution of \eqref{solvemeforvwjvminusj} down to ensure that the true valuation of $\mathcal{W}[j, v]$ is \emph{at least} $v - \Delta$, and therefore that $G[j, v]$ meets its definition.

@@

Algorithms \ref{ellisDP1}, \ref{ellisDP2}, and \ref{ellisDP3} can be written using recursive functions instead of lookup tables. However, since each function references itself \emph{twice,} the function values at each index must be recorded in the computer's memory to take advantage of overlapping indices and guarantee the runtimes promised above. A discussion of these and other implementation appears in Kapur (2022), along with a computational study demonstrating the efficiency of Algorithms \ref{ellisDP1} and \ref{ellisDP3}.

@@

This article has introduced a novel nonlinear knapsack problem called the college application problem. Our results suggest that the college application is a relatively easy instance of submodular maximization: The FPTAS presented here induces a continuum of tight, polynomial-time approximation algorithms, whereas no polynomial-time algorithm for the general problem of maximizing a nondecreasing submodular function over a knapsack constraint provides an approximation guarantee better than $1 - 1/e$.

One interpretation of our result is that the objective function of the college application, though nonlinear, has similar regularity properties to a linear function. Further evidence for this view is found in Kapur (2022)'s analysis of the college application problem with a cardinality constraint, which can be solved by a simple greedy algorithm. We conjecture that other results associated with maximizing linear functions may hold with respect to the objective discussed here, such the existence of a PTAS when maximizing over the intersection of knapsack constraints, and the optimality of the greedy algorithm when maximizing over a polymatroidal system (Chekuri and Khanna 2005; Edmonds 1971).
